---
title: "Validation Report - `r sprintf('%s (v%s)',(dcf <- read.dcf(file.path(Sys.getenv('INPUT_REPORT_PKG_DIR', '.'), 'DESCRIPTION')))[,'Package'], dcf[,'Version'])`"
date: "`r Sys.time()`"
date-format: "ddd MMM DD hh:mm:ss A YYYY"
format:
  typst:
    toc: true
    section-numbering: 1.1.1
    df-print: default
    margin:
      x: 2cm
      y: 2cm
---

```{r setup, include = FALSE}
options(width = 80L, covr.record_tests = TRUE)
remotes::install_local(
  Sys.getenv('INPUT_REPORT_PKG_DIR', '.'),
  force = TRUE,
  quiet = TRUE,
  INSTALL_opts = "--with-keep.source",
  upgrade = "never"
)
library(magrittr)
library(tinytable)
```

# Context

This report was generated via the GH-action insightsengineering/validatoR (gh-action ID: `r Sys.getenv("GITHUB_ACTION")`). It produces automated
documentation of the installation of this package on an open source R environment, focussing on:

- Installation environment description
- Testing coverage
- Traceability matrix of specifications (documented behaviours) and testing
- Risk assessment benchmarks

This report is fully automated, so is limited to assess whether unit tests and documentation are present and can execute without error. An assessment would be required that the tests and documentation are meaningful. Validation aims to be system independent as the underlying workflow is based on the ["composite"](https://docs.github.com/en/actions/creating-actions/creating-a-composite-action) type of Github Action.

# Installation environment and package

## System Info

```{r execution_info, echo = FALSE}
tt_sys_info_df <- data.frame(
  Field = c("OS", "Platform", "System", "Execution Time"),
  Value = c(
    sessionInfo()$running,
    R.version$platform,
    R.version$system,
    format(Sys.time(), tz = "UTC", usetz = TRUE)
  )) %>%
  dplyr::mutate(Value = paste0("``` ", Value, " ```"))

tt(
  tt_sys_info_df,
  caption = "System information",
  theme = "striped"
)
```

## Package installed

```{r version_control, echo = FALSE}
# find .git dir containing the package directory
gd <- system(
  sprintf("cd '%s' && git rev-parse --absolute-git-dir", Sys.getenv('INPUT_REPORT_PKG_DIR', '.')),
  intern = TRUE
)
# define reused git args to be sure we're picking up the right git info
gd <- sprintf("--git-dir='%s'", gd)
wt <- sprintf("--work-tree='%s'", Sys.getenv('INPUT_REPORT_PKG_DIR', '.'))
tt_git_df <- data.frame(
  Field = c("server", "repository", "git ref", "branch", "commit `SHA1`", "commit date"),
  Value = c(
    Sys.getenv('GITHUB_SERVER_URL'),
    Sys.getenv('GITHUB_REPOSITORY'),
    Sys.getenv('GITHUB_REF'),
    system2(
      "git",
      list(gd, wt, "rev-parse", "--abbrev-ref", "HEAD"),
      stdout = TRUE
    ),
    system2("git", list(gd, wt, "rev-parse", "HEAD"), stdout = TRUE),
    system2(
      "git",
      list(gd, wt, "show", "-s", "--format=%ci", "HEAD"),
      stdout = TRUE
    )
  )) %>%
  dplyr::mutate(Value = paste0("``` ", Value, " ```"))

tt(
  tt_git_df,
  caption = "Git information",
  theme = "striped"
)
```

## R Session Info

```{r session_info, echo = TRUE, eval = TRUE}
sessionInfo()
capabilities()
```

# Metric based risk assessment

The following metrics are derived from the `riskmetric` R package. Metrics overlapping with `covr` and `R CMD Check` are removed.

```{r riskmetric, echo = FALSE, eval = TRUE}
d_riskmetric <- Sys.getenv('INPUT_REPORT_PKG_DIR', '.') %>%
  riskmetric::pkg_ref() %>%
  riskmetric::pkg_assess() %>%
  purrr::map(1) %>%
  lapply(as.character) %>%
  tibble::enframe() %>%
  tidyr::unnest(cols = dplyr::everything()) %>%
  # add labels
  dplyr::left_join(
    lapply(riskmetric::all_assessments(), attributes) %>%
      purrr::map_df(tibble::as_tibble),
    by = c("name" = "column_name")
  )

tt_riskmetric_df <- d_riskmetric %>%
  dplyr::filter(
    name %in% c(
      "news_current", "has_vignettes",
      "license", "downloads_1yr"
    )
  ) %>%
  dplyr::select(Metric = label, Status = value)

tt(
  tt_riskmetric_df,
  caption = "Package info assessed by the R package riskmetric",
  theme = "striped"
)
```

# Installation documentation

## `R CMD check`

```{r r_cmd_check, echo = FALSE, eval = TRUE}
rcmdcheck_results <- rcmdcheck::rcmdcheck(
  Sys.getenv('INPUT_REPORT_PKG_DIR', '.'),
  args = c(
    "--timings",             # include execution times in output
    "--no-build-vignettes",  # run vignette code, but disable pdf rendering
    "--no-manual"            # disable pdf manual rendering
  ),
  quiet = TRUE
)

cat(rcmdcheck_results$stdout)
cat(rcmdcheck_results$stderr)
```

## Testing Coverage

```{r coverage, echo = FALSE, eval = TRUE}
covr_results <- covr::package_coverage(Sys.getenv('INPUT_REPORT_PKG_DIR', '.'))
covr_results
```

## Traceability

Tracebility matrix that maps each unit test to the corresponding documentation, creating a link
between intended use and testing.

```{=typst}
#show figure: set block(breakable: true)
```

### Testing matrix

```{r traceability, echo = FALSE, eval = TRUE}
if (require("covtracer", quietly = TRUE)) {
  covtracer_df <- test_trace_df(covr_results)
  covtracer_df$filename <- basename(covtracer_df$filepath)

  covtracer_df_clean <-  covtracer_df %>%
    dplyr::select(
      alias,
      test_name,
      file,
      is_exported,
      doctype,
      direct
    ) %>%
    dplyr::filter(!doctype %in% c("data", "class"))  %>%  # ignore objects without testable code
    dplyr::filter(is_exported == TRUE) %>% # Only exported
    dplyr::filter(!duplicated(.)) %>%
    dplyr::mutate(alias = paste0(alias, "()")) %>%
    dplyr::mutate(file = paste0("man/", file))

  tt_covtracer_df <- covtracer_df_clean %>%
    dplyr::filter(!duplicated(.)) %>%
    dplyr::filter(!is.na(test_name)) %>%
    dplyr::filter(!is.na(file)) %>% # unexported?
    dplyr::arrange(file) %>%
    dplyr::mutate(test_name = paste0("``` ", test_name, " ```")) %>%
    dplyr::mutate(file = paste0("``` ", file, " ```")) %>%
    dplyr::select(`Test Description` = test_name, Documentation = file)

  if (nrow(tt_covtracer_df) > 0) {
    tt(
      tt_covtracer_df,
      caption = "Tracebility matrix mapping unit tests to documented behaviours.",
      theme = "striped"
    )
  } else {
    cat("No test suites.")
  }
} else {
  cat("{covtracer} not available to produce a traceability matrix")
}
```

```{=typst}
#show figure: set block(breakable: false)
```

### Untested

```{r traceability2, echo = FALSE, eval = TRUE}
if (require("covtracer", quietly = TRUE)) {
  tt_untested_df <- covtracer_df_clean %>%
    dplyr::filter(is.na(test_name)) %>%
    dplyr::arrange(alias) %>%
    dplyr::mutate(alias = paste0("``` ", alias, " ```")) %>%
    dplyr::mutate(file = paste0("``` ", file, " ```")) %>%
    dplyr::select(`Exported package object` = alias,  Documentation = file)

  if (nrow(tt_untested_df) > 0) {
    tt(
      tt_untested_df,
      caption = "Untested behaviours: documentation that is not covered by any test.",
      theme = "striped"
    )
  }
} else {
  cat("{covtracer} not available to produce a traceability matrix")
}
```

### Testing granularity

An indicator of test granularity by whether the function is directly tested.

```{r traceability3, echo = FALSE, eval = TRUE}
if (require("covtracer", quietly = TRUE)) {
  tt_granularity_df <- covtracer_df_clean %>%
    dplyr::group_by(alias) %>%
    dplyr::summarize(any_direct_tests = any(direct, na.rm = TRUE)) %>%
    dplyr::arrange(alias) %>%
    dplyr::mutate(alias = paste0("``` ", alias, " ```")) %>%
    dplyr::mutate(any_direct_tests = paste0("``` ", any_direct_tests, " ```")) %>%
    dplyr::select(`Exported package object` = alias,  `Tested Directly` = any_direct_tests)

  tt(
    tt_granularity_df,
    caption = "Granularity of unit tests: directly tested exported functions.",
    theme = "striped"
  )
} else {
  cat("{covtracer} not available to produce a traceability matrix")
}
```
